---
title: "PCA"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

### Tutorial
https://satijalab.org/seurat/articles/pbmc3k_tutorial.html

### Hyperparam used in the paper
https://github.com/naomihabiblab/5xFAD-sNucSeq/blob/master/FUNCTIONS_SINGLE_CELL_ANALYSIS.r



### Read in the data

```{r}
library(readr)
EZ_Batch1 <- read_csv("EZ_Batch1.csv")
EZ_Batch1<-as.data.frame(EZ_Batch1)
```
```{r}
colnames(EZ_Batch1)[1] <- "gene"
rownames(EZ_Batch1) <- EZ_Batch1$gene
EZ_Batch1$gene <- NULL
head(EZ_Batch1)

```

### Createa Seurat object
```{r}
library(Seurat)
obj <- CreateSeuratObject(counts=EZ_Batch1, project="EZ_Batch1", min.cells = 5)
obj
```


## Some preprocessing work
Mitochondria (genes begin with "MT-") also has RNAs but is not in our study scope so we remove them here.
```{r}
lowT=100
HighT=Inf
obj[["percent.mito"]] <- PercentageFeatureSet(obj, pattern = "^MT-")
VlnPlot(object = obj, features = c("nFeature_RNA", "nCount_RNA", "percent.mito"), ncol = 3, pt.size = 0, group.by = "orig.ident", sort=FALSE)
obj <- subset(obj, subset = nFeature_RNA > lowT & nFeature_RNA < HighT & percent.mito < 0.05)
```
Seems like we don't have MT in our dataset. Have a look at the relationship between n_genes and N_nuclei
```{r}
FeatureScatter(obj, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
```
```{r}
Norm_PCA <- function(obj, maxPC = 20, nVar = 2000,ClusRes=0.8) {
  obj <- NormalizeData(object = obj,assay = "RNA",normalization.method = "LogNormalize")
  obj <- FindVariableFeatures(object = obj,assay = "RNA",nfeatures = nVar)
  obj <- ScaleData(object = obj, assay = "RNA",features = VariableFeatures(object = obj))
  obj <- RunPCA(object = obj,assay = "RNA",features = VariableFeatures(object = obj),npcs = 50,ndims.print = 6)
}

obj = Norm_PCA(obj)
ElbowPlot(object = obj, ndims = 50,reduction = "pca")
```

Let's find a right place to cut the PCA
```{r}
## https://hbctraining.github.io/scRNA-seq/lessons/elbow_plot_metric.html#:~:text=The%20elbow%20plot%20is%20helpful,the%20majority%20of%20the%20variation.
library("ggplot2")

Find_Elbow <- function(obj, cumu_pct_thd=95, pct_thd=3){
                       
  pct <- obj[["pca"]]@stdev / sum(obj[["pca"]]@stdev) * 100
  # Calculate cumulative percents for each PC
  cumu <- cumsum(pct)
  # Determine which PC exhibits cumulative percent greater than 90% and % variation associated with the PC as less than 5
  co1 <- which(cumu > cumu_pct_thd & pct < pct_thd)[1]
  co2 <- sort(which((pct[1:length(pct) - 1] - pct[2:length(pct)]) > 0.1), decreasing = T)[1] + 1
  pcs <- min(co1, co2)
  plot_df <- data.frame(pct = pct, 
             cumu = cumu, 
             rank = 1:length(pct))
  # Elbow plot to visualize 
  ggplot(plot_df, aes(cumu, pct, label = rank, color = rank > pcs)) + 
    geom_text() + 
    # geom_vline(xintercept = cumu_pct_thd, color = "grey") + 
    # geom_hline(yintercept = min(pct[pct > pct_thd ]), color = "grey") +
    labs(x="Cumulative percentage of variation", 
         y="Percentage of variation",
         fill = "After elbow point") + 
    theme_bw()
}
Find_Elbow(obj)
```
We'll use dimension 1-20
```{r}
maxPC = 20
ClusRes = 0.8
obj <- RunTSNE(object = obj,reduction = "pca",dims = c(1:maxPC), features = VariableFeatures(obj),tsne.method = "Rtsne")
obj <- FindNeighbors(object = obj,reduction = "pca",dims = c(1:maxPC),features = VariableFeatures(obj),k.param = 40)
obj <- FindClusters(object = obj,resolution = ClusRes)
```

```{r}
DimPlot(object = obj,reduction = "tsne", label=TRUE)
```


```{r}
obj <- RunUMAP(obj, dims = 1:maxPC)
DimPlot(obj, reduction = "umap", label=TRUE)
```

```{r}
# find all markers of each clusters
for (x in 1:21) {
  print(x)
  x_markers <- FindMarkers(obj, ident.1 = x, min.pct = 0.25,  test.use = "wilcox")
  write.csv(x_markers,sprintf("markers/c%s.csv",x)
)
}

```
## Check out our analysis in the find_marker.ipynb. We found that cluster 17 is microglia 


```{r}
c5 <- subset(obj, idents = c("5"))
```
```{r}
saveRDS(c5, file = "c5.rds")
```